Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2024-01-08 E-Hypergraphs-LICS24.tex
intro

CHRIS: CITE HASEGAWA - TRACE MONOIDAL CATEGORIES, AND PERDIX. 
DAN: I SUGGEST TO CITE HASEGAWA IN FURTHER WORK, FOR GENERALISING THE THEORY TO TRACED MC.
DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.

@@

We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$.\footnote{Note, in particular,  that semilattices do not require a unit for $+$. This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane} ---a degeneracy we wish to avoid.} More precisely our hom-sets will be given the structure of a semilattice.  
Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).  In this case,  hom-sets are generalized to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition 
\[
	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
\]
is defined as a $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.  Technically, this amounts to asking that the tensor product is an enriched functor. 
We omit further details of the general case of enrichment here, and instead work with the following more concrete defintion, where $M = \textbf{SLatt}$.

>>>>

CHRIS: CITE HASEGAWA - TRACE MONOIDAL CATEGORIES, AND PERDIX. 
DAN: THIS NEEDS TO BE CAREFULLY EXPANDED TO NOT OFFEND POTENTIAL REVIEWERS.

@@

We will use enrichment to axiomatize the e-box structure as the ``join" of two morphisms $f,g: A \to B$ as $f + g: A \to B$. Note, in particular,  that semilattices do not require a unit for $+$. This is important, since, in any commutative monoid enriched category,  if there exists a categorical product, then it must be a categorical biproduct \cite{maclane} ---a degeneracy we wish to avoid. More precisely our hom-sets will be given the structure of a semilattice.  
Generally,  we may define a category \textit{enriched} in any other monoidal category $M$ (\textit{e.g.},  \textbf{SLatt}).  In this case,  hom-sets are generalized to hom-objects of $M$ (\textit{e.g.},  a semilattice) and composition 
\[
	\circ: Hom(B,C) \otimes Hom(A,B) \to Hom(A,C)
\]
is defined as a $M$-morphism.  In other words,  composition must respect the additional structure on hom-sets.  
In our case,  the category we wish to enrich is also monoidal,  so we additionally ask that the monoidal structure also respects this structure.  Technically, this amounts to asking that the tensor product is an enriched functor. 
We omit further details of the general case of enrichment here, and instead work with the following more concrete defintion, where $M = \textbf{SLatt}$.
